---
title: "DATA 550 Final Report"
author: Helen Chen
output: html_document
date: "`r Sys.Date()`"
---

```{r, here-i-am, include = FALSE}
here::i_am(
  "Report.Rmd"
)
```

```{r, load-data, include = FALSE}
shoe_data <- read.csv(
  file = here::here("data/global_sports_footwear_sales_2018_2026.csv")
)
```

The dataset I used for my final project is from Kaggle
(https://www.kaggle.com/datasets/aliiihussain/sports-footwear-sales-and-consumer-behavior?resource=download), and it contains consumer and sales data on shoe purchases from 2018 to 2026. The author of the dataset is Ali Hussain. The dataset contains 30,000 rows and 18 columns, and the columns include the shoe brand (i.e. Nike, Adidas, Reebok, etc.), color of the shoe, country where the sale occurred, the discount applied on the sale (in %), and more. 

A question I explored using this dataset was: what are the main variables for predicting revenue (in USD)? I was interested in this as it relates to revenue forecasting in the shoe industry.

First, I did some EDA and presented a table and two boxplots to show different associations. Then, I ran a linear regression model with group LASSO for feature selection. I used group LASSO for all the categorical variables, so it wouldn’t separate the categorical levels of these variables, which is helpful for interpretability. Because it would be computationally intensive to run group LASSO for the entire dataset, I focused on data from the past five years (2022-2026). Then, I re-ran a linear regression model with all the variables that had non-zero coefficients from group LASSO. I presented the results of my final model in a well-formatted table.

# Tabular Analysis:

```{r, table-1}
table_1 <- readRDS(
  file = here::here("output/table_1.rds"))
  
table_1
```

The table above shows the total shoe units sold and total revenue (in USD) by country, in increasing order of total revenue. We can see that Pakistan has the fewest units sold and the lowest total revenue, and UAE has the greatest number of units sold and the highest total revenue. From the table, we can also see that there's a positive association between total revenue and units sold, where as units sold increases, the total revenue also increases. This is understandable, as revenue is equal to quantity * price. This also suggests that the shoe prices in each of the countries are generally similar. If the prices were very different, for example, if in UAE they had much more expensive shoes, then the units sold may not be large but the revenue could still be high.

# Plots:

```{r, boxplot1}
knitr::include_graphics(
  here::here("output/boxplot1.png")
)
```

The boxplot above shows the distribution of revenue for each of the shoe brands in the dataset, in increasing order of median revenue. We can see that the median revenues are generally similar for all the shoe brands. New Balance has the highest median revenue, and Reebok has the lowest.

```{r, boxplot2}
knitr::include_graphics(
  here::here("output/boxplot2.png")
)
```

The boxplot above shows the distribution of revenue for each of the shoe categories, in increasing order of median revenue. We can see that the median revenues are generally quite similar for all the shoe categories in the dataset. The training and lifestyle categories have the highest median revenues, and the basketball category has the lowest. This is understandable, as training and lifestyle are more generic and could apply to a larger group of people, whereas basketball shoes are more specific and may not be applicable/useful for most people.

# Group LASSO applied to linear regression model for feature selection:

I applied group LASSO to a linear regression model for feature selection. The response variable was revenue (in USD), and the predictors were: shoe brand, shoe category, gender, shoe size, shoe color, discount percentage, sales channel, units sold, payment method, country of sale, customer income level, and customer rating. I used a 70%/30% training/testing set split, 5-fold cross validation, and I set seed at '123' for reproducibility. I focused on the years 2022-2026 here, as it's computationally intensive to run group LASSO on the entire dataset, which is very large. Using cross validation, I chose the lambda that minimized the CV error, and I looked at the coefficient estimates under that lambda value. The results are presented below:

```{r, feature-selection-lasso}
coeffs_for_model <- readRDS(
  file = here::here("output/model_coeffs.rds"))
  
coeffs_for_model
```

The predictors with non-zero coefficients were: shoe brand, gender, shoe color, discount percentage, units sold, payment method, customer income level, and customer rating. 

The CV RMSE on the test set was 110.757. The maximum revenue in the test set was 876 and the minimum 42. The relative error is around 13%, which is moderate for revenue forecasting.

# Linear regression model:

I re-ran a linear regression model with the variables that had non-zero coefficients from group LASSO. The response was revenue (in USD), and the predictors were shoe brand, gender, shoe color, discount percentage, units sold, payment method, customer income level, and customer rating. The results are presented in the table below:

```{r, model1}
model1 <- readRDS(
  file = here::here("output/model1_summ.rds"))

model1
```

The variables that were statistically significant on the alpha = 0.05 level were: units sold, discount percentage on the shoe, customer income level (medium and high vs low), and if the shoes were for women (vs men's shoes).

The association between units sold and revenue aligns with what we saw in the first table. This is understandable, as revenue is equal to quantity * price. It also suggests that the shoe prices in each of the countries are generally similar. As discount percentage increases, we expect the revenue to decrease. This is understandable, as when price decreases, revenue generally also decreases (per the formula above). It also suggests that even with discounts, there isn't a larger quantity of shoes sold to offset the revenue decrease with lower prices.

Interestingly, customer income levels of medium or high compared to low were associated with lower revenues. Also, in general, women's shoes were associated with higher revenues compared to men's shoes.
